---
title: "Vessel cooperation ~ time preliminary exploration"
author: "Emily Nazario"
date: "`r Sys.Date()`"
format:
 html: 
  self-contained: true
editor: visual
toc: TRUE
toc-title: "On this page"
theme: yeti
fontcolor: "#134f5c"
---

```{r}
#| warning: false
#| include: false
#| echo: false
#| label: libraries and data

library(tidyverse)
library(lme4)
library(here)

vsr_dat <- read.csv(here("data/socal_vsr.csv"))

```

# Vessel compliance time series

Pooled data across vessel types and took the average compliance per year+month. Plotted means represent average compliance across all ships per year and month.

```{r}
#| echo: false
#| warning: false
#| label: compliance time series

vsr_dat <- vsr_dat %>% 
  mutate(date = as.Date(date, format = "%m/%d/%Y"), 
         year = year(date), 
         mo = month(date),
         yr_mo = zoo::as.yearmon(date))

vsr_dat %>% 
  group_by(yr_mo) %>%
  summarise(mean_co = mean(cooperation, na.rm = TRUE), 
            sd_co = sd(cooperation, na.rm = TRUE)) %>%
ggplot(aes(yr_mo, mean_co)) +
  geom_point(size = 3)+
  tidyquant::theme_tq()+
  geom_path(size = 1, alpha = 0.5)+
  xlab("")+
  ylab("Mean monthly cooperation (%)")+
  theme(legend.position = "none", 
        axis.title.y = element_text(size = 16, color = "black"), 
        axis.text = element_text(size = 14, color = "black"))
```

# Vessel compliance time series w/ points per ship type

Pooled data across ship types and year. Black plotted points represent the mean annual cooperation rate across all ship types, while the grey points are the mean annual cooperation rate per ship type.

```{r}
#| echo: false
#| warning: false
#| label: compliance time series per year w/ ship type

vsr_dat_ship <- vsr_dat %>%
  group_by(year, ship_type) %>%
  summarise(mean_co = mean(cooperation, na.rm = TRUE), 
            sd_co = sd(cooperation, na.rm = TRUE))

vsr_dat %>% 
  group_by(year) %>%
  summarise(mean_co = mean(cooperation, na.rm = TRUE), 
            sd_co = sd(cooperation, na.rm = TRUE)) %>%
ggplot(aes(year, mean_co)) +
  geom_point(data = vsr_dat_ship, aes(year, mean_co), color = "grey")+
  geom_point(size = 3)+
  tidyquant::theme_tq()+
  geom_path(size = 1, alpha = 0.5)+
  xlab("")+
  ylab("Mean annual cooperation (%)") +
  theme(legend.position = "none", 
        axis.title.y = element_text(size = 16, color = "black"), 
        axis.text = element_text(size = 14, color = "black"))

```

# Vessel compliance \~ time GLMMs

## Wilcox normality test

```{r}
#| warning: false

wilcox.test(vsr_dat$cooperation)

```

::: panel-tabset
## Vessel compliance \~ year_month

```{r}
#| warning: false
#| label: compliance ~ yr_mo glmm

comp_time_yr_mo <- glmer(cooperation/100 ~ yr_mo + (1|ship_type), data = vsr_dat, family = "binomial");summary(comp_time_yr_mo)

plot(comp_time_yr_mo)
lattice::qqmath(comp_time_yr_mo)

```

## Vessel compliance \~ year

```{r}
#| warning: false
#| label: compliance ~ year glmm

comp_time_yr <- glmer(cooperation/100 ~ year + (1|ship_type), data = vsr_dat, family = "binomial");summary(comp_time_yr)

plot(comp_time_yr)
lattice::qqmath(comp_time_yr)

```

## Vessel compliance \~ month

```{r}
#| warning: false
#| label: compliance ~ month glmm

comp_time_mo <- glmer(cooperation/100 ~ mo + (1|ship_type), data = vsr_dat, family = "binomial");summary(comp_time_mo)

plot(comp_time_mo)
lattice::qqmath(comp_time_mo)

```
:::
